#Case Study : Customer Segmentation
```{r}
setwd("C:/Users/Nupur Rathi/Projects/Customer Segmentation/customer-segmentation-dataset")
cust<-read.csv("Mall_Customers.csv")

```
#Understanding Data
```{r}
head(cust)
dim(cust)
names(cust)

#Renaming the columns
library(dplyr)
cust<-cust%>%
  rename("Spend"=Spending.Score..1.100.,"Income"=Annual.Income..k..)

#Checking structure
str(cust)
#structure or data types are in correct format

#Duplicate obs check
anyDuplicated(cust)

#Missing value check
colSums(is.na(cust))


```

#Data visualization
#Gender
```{r}
gender<-table(cust$Gender)
barplot(gender,xlab="Gender",ylab = "Count",main = "Gender Comparision",col = c("pink","blue"),legend=rownames(gender))


```

```{r}
pct=gender/sum(gender)*100
lbs<-paste(c("Female","Male"), pct, "%")
lbs
library(plotrix)
pie3D(gender, labels = lbs)
```

#Age
```{r}
hist(cust$Age,col = "grey",xlab = "Customer Age",ylab="Count",main="Distribution of CustomerAge")
boxplot(cust$Age,col="pink")
```

#Income
```{r}
summary(cust$Income)
boxplot(cust$Income)
hist(cust$Income,col="blue",xlab = "Customer Income",ylab="Count" , main="Distribution of Income")
```
#Min Income  : 15 , MaxIncome is 137, average Income : 60 k , median : 70 k

#Spend
```{r}
summary(cust$Spend)
boxplot(cust$Spend,col="maroon")
hist(cust$Spend,col="purple",labels = TRUE)
```

#More ppl spend 40-50 k

#Clustering

#Heirarchial Clustering
```{r}
d_euc<-dist(cust,method = "euclidean")
print(d_euc)
hclust1<-hclust(d_euc,method = "average")
plot(hclust1,labels = cust$CustomerID)
rect.hclust(hclust1,k=8,border = "blue")

#Adding Column clusters
cust$clusters<-cutree(hclust1,k=8)
View(cust)

#profiling the clusters
aggr<-aggregate(cust[,-c(1:2)],list(cust$clusters),mean)
aggr

clus.profile<-data.frame(cluster=aggr[,5],
                         frequency=as.vector(table(cust$clusters)),
                         aggr[,-c(1,5)])
clus.profile
```



#Scale Data: In this case we dont need scale but normally when we have uncleaned data income in thousands and age in tens, we need to do scaling.
```{r}
custScale<-scale(cust[,-c(1:2)])  # By Default - Z Scaling
head(cust)
```

#k-means
#Identifying the optimal number of clusters
```{r}
install.packages("NbClust")
install.packages("factoextra")
library(NbClust)
library(factoextra)

#To have a plot of optimal clusters.
fviz_nbclust(custScale,kmeans,method = "wss")+
   geom_vline(xintercept = 5, linetype = 2)+
  labs(subtitle = "Elbow method")

```

#kmeans clustering
```{r}
model1<-kmeans(x=custScale,centers=4)
model1
model2<-kmeans(x=custScale,centers = 5)
model2
```

```{r}
cust$kc4<-model1$cluster
cust$kc5<-model2$cluster

write.csv(cust,"cust.csv")
```

```{r}

ggplot(cust,aes(x=Income,y=Spend))+
  geom_point(aes(color=as.factor(kc5)))+
  scale_color_discrete(labels=c("Cluster 1","Cluster 2","Cluster 3","Cluster 4",
                                "Cluster 5"),name="Clusters")
  
```

